---
title: "Usage of VL+eBayes+ash"
output: html_document
---

Here we illustrate the usage of our proposed method *VL+eBayes+ash* on a simulated toy dataset. The R packages [limma](https://bioconductor.org/packages/release/bioc/html/limma.html), [edgeR](https://bioconductor.org/packages/release/bioc/html/edgeR.html) and [ashr](https://cran.r-project.org/web/packages/ashr/index.html) are required.

First we simulate an RNA-seq dataset (1000 genes, 2 samples in group A and 2 samples in group B). 

```{r}
set.seed(1234)

# Simulate a ngene*nsamp count matrix
# rows are genes and columns are samples
ngene = 1000
nsamp = 4
condition = c("A","A","B","B")  # group indicator
data = matrix(rpois(ngene*nsamp, 500), nrow=ngene)
colnames(data) = condition
head(data)
```

Now we estimate the "effect size" (i.e. the difference in expression levels between the two groups for each gene) using our proposed method *VL+eBayes+ash*. 

```{r,message=FALSE,warning=FALSE}
library(limma)
library(edgeR)
library(ashr)

# Calculate normalization factors
dgecounts = calcNormFactors(DGEList(counts=data, group=condition))
  
# Voom transformation
design = model.matrix(~condition)
v = voom(dgecounts,design,plot=FALSE)
  
# Use lmFit to estimate effect sizes betahat
lim = lmFit(v)

# Then use eBayes to shrink the standard errors
lim = eBayes(lim)

# Get betahat, the shrunk s.e. and moderated d.f
betahat = lim$coefficients[,2]
se = lim$stdev.unscaled[,2]*sqrt(lim$s2.post) # EB shrunk s.e.
df = lim$df.total[1]
  
# Fit ash with the shrunk s.e. and moderated d.f
fit = ash(betahat, se, df=df)
```

After getting the ashr fitted object `fit`, we can fetch the estimated effect sizes and q-values for each gene by the following commands. 

```{r}
# Gene-wise shrunk estimates for the effect sizes are
beta_est = fit$result$betahat
summary(beta_est)

# Gene-wise q-values for testing if true effect is 0
qval = fit$result$qvalue
summary(qval)
```

From the simulation setting, we know this is a pure null dataset (all effect sizes should be 0). The above results show that the estimated effect sizes are very small around 0 with big q-values, which suggests they are null genes.  